

```{r include=FALSE}
library(BookUtilities)
source(system.file("DataComputing-Settings.R", package = "BookUtilities"), local=TRUE)
showAnswers <- FALSE
```

The `OrdwayBirds` data table is a historical record of birds captured and released at the Katharine Ordway Natural History Study Area, \index{Ordway!Natural History Study Area} a 278-acre preserve in Inver Grove Heights, Minnesota, owned and managed by Macalester College.\index{MacalesterCollege}  Originally written by hand in a field notebook, the entries have been transcribed into electronic format under the supervision of Jerald Dosch, \index{Dosch!Jerald} Dept. of Biology, Macalester College.


Due to mistakes in data entry, the `SpeciesName` variable needs some fixing.  `SpeciesName` is intended to identify the species of each of the birds, but the  spelling often varies among birds of the same biological species.  This leads to mis-classification of birds.  There are also problems with the `Month` and `Day` variables; they are supposed to be numerical, but mistakes prevent them from being correctly identified as such.

Fortunately, all these errors are easy to correct.  The data table `OrdwaySpeciesNames` collects together all the variant spellings.  Entry by entry, each mis-spelling was translated (by a human) into a standardized spelling.  Thus, `join()` can be used to correct the mis-spellings in the `OrdwayBirds` table.

You are going to look at the month-to-month presence of different species.  Think of your assignment as creating a manual for birders to guide them to the correct time of year to visit Ordway to see a particular species.

## Some simple data cleaning

There are many variables that you won't need for this activity, and you still have to fix the `Month` and `Day` variables.  To keep things simple, cut and paste this command into a chunk at the start of the document.

## **Step 0** Get the data table

```{r}
OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month=as.numeric(as.character(Month)), 
          Day=as.numeric(as.character(Day)))
```



The `mutate()` step is part of the data cleaning process, arranging `Month` and `Day` as numerical variables as originally intended by the folks entering the data. 

#### **Task 1**. 

Including mis-spellings, how many different species are there in the `OrdwayBirds` data? 

```{r include=showAnswers}
OrdwayBirds %>% 
  summarise( count=n_distinct(SpeciesName))
```

Make a data table that gives the number of distinct species in the `SpeciesNameCleaned` variable in `OrdwaySpeciesNames`.  

```{r include=showAnswers}
OrdwaySpeciesNames %>%
  summarise(count = n_distinct(SpeciesNameCleaned))
```

\noindent You will find it helpful to use `n_distinct()` a reduction function, which counts the number of unique values in a variable.

#### **Task 2**

Use the `OrdwaySpeciesNames` table to create a new data table that corrects the mis-spellings in `SpeciesNames`.  This can be done easily using the `inner_join()` data verb.  


```{r}
Corrected <-
  OrdwayBirds %>%
  inner_join( OrdwaySpeciesNames ) %>% 
  select( Species=SpeciesNameCleaned, Month, Day ) %>%
  na.omit() ## cleaned up the missing ones
```


Look at the names of the variables in `OrdwaySpeciesNames` and `OrdwayBirds`.  

* Which variable(s) was used for matching cases.
* What were the variable(s) that will be added .

#### **Task 3**

Count how many bird captures there are of each of the (corrected) species?  You can call the variable that contains the count `count`.  Arrange this into descending order from the species with the most birds, and look through the list. \marginnote{Hint: Remember \texttt{n()}.  Also, one of the arguments to one of the data verbs will be \texttt{desc(count)} to arrange the cases into descending order.  Display the top 10 species in terms of the number of bird captures.}

```{r include=showAnswers}
Corrected %>%
  group_by(Species) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Define for yourself a "major species" as a species with more than a particular threshold count.  Set your threshold so that there are 5 or 6 species designated a major. 

Filter to produce a data table with only the birds that belong to a major species.\marginnote{Hint: Remember that summary functions can be used case-by-case when filtering or mutating a data table that has been grouped.}  Save the output in a table called `Majors`.

#### **Task 4**

When you have correctly produced `Majors`, write a command that produces the month-by-month count of each of the major species. Call this table `ByMonth`.

Display this month-by-month count with a bar chart arranged in a way that you think tells the story of what time of year the various species appear. You can use `barGraphHelper()` to explore different possibilities.  Use the "Show Expression" button in `mBar()` to create an expression that you can cut and paste into a chunk in your Rmd document, so that the graph gets created when you compile it.\marginnote{Warning: \texttt{barGraphHelper()} should never be a statement in your Rmd file, it needs to be used interactively from the console.}

Once you have the graph, use it to answer these questions:

1. Which species are present year-round?
#. Which species are migratory, that is, primarily present in one or two seasons?
#. What is the peak month for each major species?  
#. Which major species that are seen in good numbers for at least 6 months of the year? (Hint: `n_distinct()` and `>= 6`.) 


```{r eval=FALSE,echo=FALSE,message=FALSE,results='hide'}
OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month=as.numeric(as.character(Month)), 
          Day=as.numeric(as.character(Day)))

Corrected <-
  OrdwayBirds %>%
  inner_join( OrdwaySpeciesNames ) %>% 
  select( Species=SpeciesNameCleaned, Month, Day ) %>%
  na.omit() ## cleaned 
MaxMonth <- Corrected %>%
  group_by( Species, Month ) %>%
  summarise( n=n() ) %>% 
  filter( rank( desc(n)) < 2 )
MajorSpeciesIDs <-
  Corrected %>%
  group_by( Species ) %>% 
  summarise( n=n() ) %>%
  filter( rank( desc(n)) < 10 )
OnlyMajorSpecies <-
  Corrected %>% 
  inner_join( MajorSpeciesIDs %>% select(Species)) %>%
  droplevels( ) # drop the non-major species levels
# Or, instead of the above
OnlyMajorSpecies2 <-
  Corrected %>%
  group_by( Species ) %>%
  filter( n() > 500 ) # or whatever the threshold is.
MajorCount <- 
  OnlyMajorSpecies2 %>% 
  group_by( Species, Month ) %>% 
  summarise( count=n() )
BigMonths <-
  OnlyMajorSpecies %>%
  group_by( Month, Species ) %>%
  summarise( n=n() ) %>%
  filter( n > 0.02*sum(n) )
SixMonthsOrMore <-
  BigMonths %>% 
  group_by( Species ) %>%
  summarise( months=n() ) %>% 
  filter( months >= 6 )
TimeOfMonth <-
  OnlyMajorSpecies %>% 
  inner_join( select(MaxMonth,-n)) %>%
  mutate( Day=as.numeric(Day) ) %>%
  group_by( Month, Species ) %>%
  summarise( early=mean(Day<=10),
             mid = mean(Day>10 & Day<20),
             late= mean(Day>=20))

ByMonth <- MajorCount
p <- ggplot(data=ByMonth,aes(x=as.numeric(as.character(Month)),y=count,order=reorder(Species,count),fill=Species)) +
  geom_bar(stat='identity',position=position_dodge(width=.9)) + facet_wrap(~Species,ncol=3)
```

