---
title: "Visualizing Movie Ratings"
author: "Data Computing"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
---
```{r include=FALSE}
library(DataComputing)
library(tidyr)
library(ggplot2)
library(mosaic)
knitr::opts_chunk$set(tidy = FALSE)
```

<!-- # Read the data --- just for the draft -->

```{r eval = FALSE, echo = FALSE}
data_dir <- "~/Downloads/ml-100k/"
read_ml_file <- function(fname, data_dir, vnames, sep = "|") {
  Res <- read.csv(paste0(data_dir, fname), stringsAsFactors = FALSE, header = FALSE, sep = sep, colClasses = "character")
  names(Res) <- vnames
  Res
}
Users <-
  read_ml_file("u.user", data_dir,
                      vnames = c("user_id", "age", "sex", "occupation", "zip_code")) %>%
  mutate(age = tidyr::extract_numeric(age))

Movies <-
  read_ml_file("u.item", data_dir, 
                vnames = c("movie_id", "movie_title", "release_date", "video_release_date", "IMDb_URL", "unknown", 
                           "Action", "Adventure", "Animation", "Children's", "Comedy",  "Crime",  "Documentary",  "Drama",  "Fantasy", 
                           "Film-Noir",  "Horror",  "Musical",  "Mystery",  "Romance",  "Sci-Fi", "Thriller",  "War",  "Western"))
Movies <-
  Movies %>%
  select( - video_release_date) %>%
  mutate(release_date = lubridate::dmy(release_date)) #, video_release_date = lubridate::dmy(video_release_date))
for (k in 5:23) {
  Movies[[k]] <- as.logical(as.numeric(Movies[[k]]))
}

Ratings <-
  read_ml_file("u.data", data_dir, sep = "\t",
                vnames = c("user_id", "movie_id", "rating", "time_stamp"))
Ratings <-
  Ratings %>%
  mutate(rating = tidyr::extract_numeric(rating),
         time_stamp = as.POSIXct(as.numeric(as.character(time_stamp)), origin="1970-01-01", tz = ""))

save(Users, Movies, Ratings, file = "MovieLens.rda")
```

A set of 100,000 ratings of movies by individuals was collected in the late 1990s by the *grouplens* research team at the University of Minnesota. The grouplens team provides the data directly at <http://grouplens.org/datasets/movielens/100k/>.  These data were reformatted by for the *Data Computing* book. Downloaded them to your own computer with this statement:
```{r eval = FALSE, message = FALSE}
download.file("http://tiny.cc/dcf/MovieLens.rda", 
              destfile = "MovieLens.rda")
```

You only need to download the data once. But each time you start a new R session^[Every time you *knit* a document, you are starting a new session just for the purpose of compiling that document.] you will need to `load()` the data to your R session. 
    
```{r echo = FALSE}
load("MovieLens.rda")
```

`MovieLens.rda` contains three data tables:

* `Ratings` has the individual movie ratings and the time at which they were entered. It also includes an ID variable for both the user and the movie.
* `Movies` provides the name of the movie and information about genres.
* `Users` gives basic information about the person who made the rating.

Your task: Construct each of these graphics.

## Show the appeal of different genres to the different sexes
```{r echo = FALSE, warning = FALSE, message = FALSE}
With_genres <-
  Movies %>% 
  tidyr::gather(key = "genre", value = "tmp", 5:23) %>%
  filter(tmp) %>%
  select(-tmp) 
All <- 
  Ratings %>% 
  left_join(With_genres) %>%
  left_join(Users)
Tmp <- 
  All %>%
  group_by(genre) %>%
  filter(n() > 1000) %>%
  group_by(genre, sex) %>%
  summarise(mean = mean(rating), sd = sd(rating) / sqrt(n())) %>% 
  ungroup()
Tmp %>% select(-sd) %>%
  tidyr::spread(key = sex, mean) %>%
  mutate(diff = F - M)  -> Tmp2
Tmp %>%
  left_join(Tmp2) %>%
  mutate(genre = reorder(genre, desc(F))) %>%
  ggplot(aes(x = genre, y = mean, color = sex)) + 
    geom_errorbar(aes(ymin = mean - 2*sd, ymax = mean + 2 * sd), position = position_dodge(width = 0.2)) + 
    geom_point(position = position_dodge(width = 0.2)) +
  coord_flip()
```

## Which genres are related?

Look at correlation between genres `cor()`.

Another possibility, look at the co-occurence, fraction of movie genre A that is also movie genre B.
```{r}
co_occurance <- function(genres){
  f <- function(x,y) {sum(x * y) / sum(x)}
  M <- matrix(0, nrow = ncol(genres), ncol = ncol(genres))
  for (first in 1:ncol(genres)) {
    for (second in 1:ncol(genres)) {
      M[first, second] <- f(genres[[first]], genres[[second]])
    }
  }
  M <- as.data.frame(M)
  names(M) <- names(genres)
  M$genre <- names(genres)
  M %>%
    tidyr::gather(key = genre2, value = co_occur, -genre)
}
```

```{r}
Genres <- Movies[,6:23]
tmp <- cor(Genres) %>% as.data.frame(stringsAsFactors = FALSE)
tmp$genre <- row.names(tmp)
Genre_pairs <-
  tmp %>% 
  gather(key = genre2, value = correlation, -genre) %>%
  filter(genre != genre2) %>%
  group_by(genre) %>%
  mutate(cor_sign = as.character(sign(correlation))) %>%
  mutate(cor_size = abs(correlation))
Genre_co_occur <- co_occurance(Genres) %>%
  filter(genre != genre2)
Genre_pairs %>%
  ggplot(aes(x = genre2, y = genre)) + 
  geom_point(aes(size = cor_size, color = cor_sign)) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
Genre_co_occur %>%
  ggplot(aes(x = genre2, y = genre)) +
  geom_point(aes(size = co_occur)) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```

As a network
```{r warning = FALSE, message = FALSE}
library(igraph)
Keep_pairs <- 
  Genre_pairs %>%
  filter(cor_size > 0.05, cor_sign == "1") %>%
  filter(genre > genre2) 
Vertices <- Keep_pairs %>% 
  edgesToVertices(from = genre, to = genre2) 
 
Edges <- 
  Vertices %>%
  edgesForPlotting(ID = ID, x, y, Edges = Keep_pairs, from = genre, to = genre2)
Vertices %>%
  ggplot(aes(x = x, y = y)) + geom_point()+
  geom_segment(data = Edges, 
               aes(x = x, y = y, xend = xend, yend = yend, 
                   color = correlation, size = correlation)) + 
  theme_map() + 
    geom_label(aes(label = ID), fill = "white") 
```

And for co-occurances ...

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(igraph)
Keep_pairs <- 
  Genre_co_occur %>%
  filter(co_occur > 0.2) %>%
  filter(genre > genre2) 
Vertices <- Keep_pairs %>% 
  edgesToVertices(from = genre, to = genre2) 
 
Edges <- 
  Vertices %>%
  edgesForPlotting(ID = ID, x, y, Edges = Keep_pairs, from = genre, to = genre2)
Vertices %>%
  ggplot(aes(x = x, y = y)) + geom_point()+
  geom_segment(data = Edges, 
               aes(x = x, y = y, xend = xend, yend = yend, 
                   size = co_occur, color = co_occur)) + 
  theme_map() + 
    geom_label(aes(label = ID), fill = "white") 
```

## Who are the reviewers?

```{r}
Users %>%
  ggplot(aes(x = age)) + 
  geom_density(aes(fill = occupation), 
               color = NA, alpha = .7, position = "fill") + 
  facet_wrap( ~ sex)
```

```{r}
Users %>%
  ggplot(aes(x = age)) + 
  geom_density(aes(fill = sex), 
               color = NA, alpha = .4, position = "fill")
```

```{r}
Users %>%
  group_by(occupation) %>%
  tally() %>%
  arrange(desc(n))
```

## Ratings as people age

```{r fig.width=10, fig.height = 15}
All %>%
  filter( genre != "unknown") %>%
  ggplot(aes(x = age, color = sex, y = rating)) + 
  geom_smooth() + 
  facet_wrap( ~ genre, scales = "free")
```

```{r}
All %>% 
  ggplot(aes(x = age, color = sex, y = rating)) +
  geom_smooth()
```